---
title: "Bacterial DNA versus cortical content from surface remains"
author: "Alex Emmons"
date: "20 Feb 2017"
output:
  html_notebook: default
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---


##Introduction: 
The data used in the following code is metadata associated with microbial 16S rRNA and 18S rRNA sequence data taken from mutliple bone types across three human individuals that decomposed in a natural surface environment. The purpose of this script was to look at the relationship between bacterial DNA abundance data as a function of cortical bone content. 


##Results: 
```{r}
#Used to load ggplot2 to create a plot visualizing bone cortical content versus bacterial DNA quantity
library(ggplot2)


#Reads in the data and assigns it to an object metadataBone3 
metadataBone3 <- read.csv("C:/Users/Alexandra Emmons/Dropbox/BoneSurfaceProject/BoneSurface_metadata_all.csv")
               

#Create a scatter plot of bacterial DNA versus cortical bone content
graph2 <- ggplot(metadataBone3, aes(X..Cortical, Bacterial.DNA, colour=Body.region)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, inherit.aes = FALSE, aes(X..Cortical, Bacterial.DNA))

#Corrects labels and print the graphical output
graph2 + ylab("Bacterial DNA yield (copy#/ul)") + xlab("Percent Cortical Bone (%)")+ 
  scale_colour_discrete(name = "Bone Region")


```
Though bones with the highest cortical content corresponded with less bacterial infiltration, the overall relationship did not appear to be linear. 

A linear model can be used to test the relationship between bacterial DNA and cortical bone content. 
```{r}
#Builds the linear model and assigns it to object fit1
fit1 <- lm(Bacterial.DNA ~ X..Cortical, data = metadataBone3, na.action = na.exclude)

#Summarizes the data
summary(fit1)

```
The linear relationship appears to be significant; however, the model is weak (r sqaured= 0.1229)

##Discussion:
Bacteiral DNA infiltration is not necessarily controlled by cortical bone content; though, the two seem to be related. 


